Reliable feature point matching without the use of geometric constraints

### 2013-06-07 text/clustering.tex
V 2.6

\documentclass[12pt,journal]{IEEEtran}
%\documentclass{article}
\usepackage{footnote}
\usepackage[font={small}]{caption}
\usepackage{sidecap}
\usepackage{graphicx}
\usepackage{cite}
\usepackage{hyperref}
\usepackage[stable]{footmisc}
\usepackage{caption}
\usepackage{adjustbox}
\usepackage{subcaption}
\usepackage{algpseudocode}
\usepackage{algorithm}
\usepackage{amsmath}
\usepackage{amssymb}

@@

where we don't know if two images have any corresponding areas or 

@@

constraints used to evaluate the matches such as angular constraints 

@@

%proposes clustering feature points by using their geometric position 
%and exclude correspondences that are geometrically deviant. A more 
%sophisticated approach is introduced in \cite{wu2011robust} where the 
%areas are created by using the MSER feature detector to find areas of 
%interest in the images. These areas can be defined as an ellipse with a 
%certain radius around the interest points. The feature points that fall 
%within this ellipse are then matched according to epipolar constraints.  
In practice however, there are many situations where using the geometry 
of the image to filter correspondences is not possible, such as object 
recognition where adjecent objects in one photo might be separated in 
another. Another example is images that do not overlap where algorithms 
might still find matches adhering to the imposed geometric constraints.
In addition methods using geometric constraints all require a set of 
correspondences to begin with. If this set of correspondences is 
narrowed down to a smaller set with a higher ratio of correct matches 
the result of the geometric matching will be both faster and more 
accurate.  Finally some use cases might require a performance that can't 
be achieved by a more complex geometric method where simple non 
geometric methods might be able to provide sufficient matching accuracy 
with a smaller performance penalty.

@@

In turn the two methods proposed in this paper are inspired by a simple 
but novel idea. If we have two images and a given feature point in the 
first image is better matched with other feature points from the 
\emph{same} image than points in the other image, then any matches of 
this feature point to points in the other image is deemed unreliable and 
discarded.  This approach carries no implicit assumptions about the 
geometric consistency of matches and can as such easily be extended with 
other geometric solutions when appropriate.

@@

%compare two images that don't correspond. For any proposed 

@@

	\State $f_m,f_n \gets get2NNs(f_i, features ~ \backslash ~ 

@@

In figure \ref{fig:comparemirror} an example of the result of \emph{MM} 
is illustrated. Notice how the by the bottom of the image admits 
correspondences that can conveniently be removed as done in sub figure 
\ref{fig:without} where only the actual correspondences returned by the 
algorithm are shown.

\subsection{Mirror Match with Clustering (\emph{MMC})}
%
As opposed to \emph{MM}, \emph{MMC} diverges from traditional 
non-geometric feature matching by clustering the feature points by 
similarity. This process yields partitions of fairly similar feature 
points that we can match using the same approach as \emph{MM}. 
%
%Before we introduce the implementation details we will go over the 
%problem of graph clustering and how it relates to feature matching.
%
%\subsubsection{Graph Clustering}
%
%Given a set of feature points from two images, $im_1$ and $im_2$: $F_k = 
%\{k_i, k_j$ for $k_i \in im_1, k_j \in im_2\}$ as well as a matching 
%function $M(k_i, k_j) \rightarrow \mathbb{R}$ that takes two feature 
%points in an image and returns their matching score, we can define a 
%matrix $A$ where each element $A_ij = M(k_i, k_j)$. A can be interpreted 
%as the \emph{adjecency matrix} of the fully connected graph where each 
%vertex corresponds to a keypoint and the edge between two vertices has a 
%weight equal to the distance between the two corresponding keypoints.
%
%This representation reduces the problem of partitioning the keypoints 
%into groups to that of graph clustering or community structure depending 
%on the context. In the literature there are various ways of clustering a 
%graph according to different measures of what constitutes an optimal 
%partitioning. Traditionally the most used clustering algorithms have 
%been K-means and spectral clustering, but in recent years a host of new 
%algorithms have been proposed based on both Newman's concept of graph 
%modularity\footnote{Introduced in \cite{girvan2002}, discussed in 
%\cite{brandes2007} and used in \cite{blondel2008} as well as others} as 
%well as information theoretical measures\footnote{See for example 
%\cite{rosvall2008}} and the Potts spin model from physics\footnote{Used 
%in \cite{ronhovde2009}} just to mention a few approaches. Many of the 
%new algorithms differ from K-means clustering and Spectral clustering in 
%that they don't require the number of expected to clusters to be 
%specified beforehand\footnote{Among the aforementioned methods, this is 
%true for \cite{blondel2008} and \cite{rosvall2008}}.  Furthermore, on 
%tests done using randomly generated graphs with a known partitioning 
%\cite{blondel2008}, \cite{rosvall2008} and \cite{ronhovde2009} perform 
%markedly better than spectral clustering and 
%K-means\cite{lancichinetti2009}.
%
%The performance of clustering algorithms is a complicated issue, since 
%an optimal clustering given the same graph can vary depending on the 
%application. Spectral clustering for example will usually return a 
%partitioning where each partition is roughly equal in size\footnote{As 
%mentioned in \cite{von2007}} while the Louvain 
%clustering\cite{blondel2008} might return partitions of very uneven 
%size, even if the modularity measurement has been shown to penalize very 
%small clusters\cite{brandes2007}. Both behaviours can be beneficial 
%depending on the application, but when clustering feature points, 
%maintaining clusters of an even size usually means that some clusters 
%will be '\emph{catch-all}' clusters where the feature points that don't 
%fit anywhere else are grouped together. The necessity of specifying the 
%amount of partitions in for example Spectral clustering or Pott's model 
%clustering further exacerbates the issue since smaller partitions are 
%then combined into one to achieve the right amount of partitions.

@@

While the louvain clustering algorithm doesn't require any parameters in 
itself, it tends towards clustering all feature points together in the 
same partition if the graph is very connected.  To ensure that the graph 
is well clustered, the adjacency matrix is pruned so only edges above a 
certain threshold is kept. From empirical analysis keeping around 2.5\% 
of edges seems to work well. Figure \ref{fig:graph} shows the result of 
clustering the feature points displayed as a graph.
%
\begin{SCfigure*}
	\makebox[0.7\textwidth][c]{%
		\begin{subfigure}[t]{0.28\textwidth}
			\centering
			\includegraphics[width=\textwidth]{images/MMC_graph_full}
			\caption{Full Graph}
			\label{fig:full_graph}
		\end{subfigure}%
		~ %add desired spacing between images, e. g. ~, \quad, \qquad		  
		%(or a blank line to force the subfigure onto a new line)
		\begin{subfigure}[t]{0.28\textwidth}
			\centering
			\includegraphics[width=\textwidth]{images/MMC_graph_cropped}
			\caption{Cropped Section of Graph}
			\label{fig:cropped_graph}
		\end{subfigure}%
	}%
	\caption{The partitioned feature graph. Each vertex represents a 
		feature point while an edge represent a high similarity between 
		points. A partition is a connected group marked by a color.  Due 
		to a limited amount of colors some disconnected partitions have 
		been assigned the same color. The edge color of each node 
		signifies which image it belongs to. In the detailed crop to the 
		right the various cluster sizes can be seen, ranging from 
	hundreds of feature points to only two or three.}
	\label{fig:graph}
\end{SCfigure*}

@@

The matching for \emph{MMC} \emph{getMatches} finds matches within all 
partitions with more than two elements as done in \emph{MM}. However as 
can be seen in the example in figure \ref{fig:cropped_graph} many of the 
partitions contain only two feature points from different images linked 
by one edge. In this case we compare the similarity of this match to the 
next best match across all partitions and filter the matches based on 
this ratio like in \emph{SIFT}.
%\begin{algorithm}
%\caption{Impl. of getMatches (\emph{from MMC algorithm})}
%\label{alg-getmatches}
%\begin{algorithmic}
%\Require $p$ : set of features, $t\in \mathbb{R}$, $features$ : Set of 
%all features
%\State $matches \gets \varnothing$
%\State $edges \gets \left\{getSimilarity(f_i, f_j) \mid getImg(f_i) \neq 
%getImg(f_j) \wedge f_i, f_j \in p \right\}$
%\If{$\left\vert edges \right\vert = 1$} \Comment If $\exists$ one edge 
%between images
%	\State $f_i,f_j \gets getInterImageFeatures(edges_{inter}, p)$
%	\State $f_m,f_n \gets getTwoNearestNeighbors(f_i, features ~ 
%\backslash ~ \left\{f_i\right\})$
%	\State $f_s,f_t \gets getTwoNearestNeighbors(f_j, features ~ 
%\backslash ~ \left\{f_j\right\})$
%	\State $ratio_i \gets getSimilarity(f_i, f_n) / getSimilarity(f_i, 
%f_j)$
%	\State $ratio_j \gets getSimilarity(f_j, f_t) / getSimilarity(f_j, 
%f_i)$
%	\If{$ratio_i < t\wedge ratio_j < t$}
%		\State $matches \gets matches \cup (f_i, f_j)$
%	\EndIf
%\ElsIf{$\left\vert edges \right\vert > 1$} \Comment If $\exists$ more 
%than one edge between images
%	\ForAll{$f_i \in p$}
%		\State $f_m,f_n \gets getTwoNearestNeighbors(f_i, p ~ \backslash 
%~ \left\{f_i\right\})$
%		\State $ratio \gets getSimilarity(f_i, f_n) / getSimilarity(f_i, 
%f_m)$
%		\If{$getImg(f_i) \neq getImg(f_m) \wedge ratio < t
%\wedge (f_m, f_i) \not\in matches$}
%			\State $matches \gets matches \cup (f_i, f_m)$
%		\EndIf
%	\EndFor
%\EndIf
%
%\Return matches
%\end{algorithmic}
%\end{algorithm}
%
\section{Experiments}
\label{experiment}
%
To reliably measure the accuracy of a matching method on real images we 
either need a set of image pairs tied by a homography or manually count 
the number of inliers. The latter becomes unfeasible the moment we 
attempt to test a non-trivial set of images.

In \cite{mikolajczyk2005performance} Mikolajczyk and Schmid introduces a 
set of images to compare the performance of feature detectors. The test 
set covers different types of image variance such as different lighting, 
blur, rotation as well as viewpoint change. Inspired by this 
dataset\footnote{In particular by the 'Graf' image set} and motivated by 
the need for more image pairs featuring viewport changes, we have 
compiled a set of $8$ image pairs consisting of the same subject taken 
from two different angles. The images are collected from Flickr's 
database of images published under a creative commons 
licence\footnote{\href{http://creativecommons.org/}{www.creativecommons.org}, 
a copyleft license permitting redistribution} and feature murals which 
makes it possible to relate points in the image pairs with a homography.  
The images have been cropped to feature the same motive and resized to 
$900$px by $600$px. This dataset will be referred to as the 
\emph{Murals} dataset in this paper. The images can be seen in figure 
\ref{fig:murals}.

@@

%Inspired by the graffiti image we have collected a set of image pairs 
%featuring murals of a few different artists\footnote{Including Banksy, 
%Blu and Shepard Fairey among others}. Each pair consist of images of 
%the same motive taken from different angles, often including repetitive 
%patterns and texture. Based on this image set the two algorithms are 
%tested on a series of image patches from each pair. To verify that the 
%algorithms proposed are reliable we need to test that we get good 
%matches when images overlap and that we get no matches when they don't.  
%This means that it isn't enough just testing against pairs of images 
%that we know match.  We also need to test against pairs that might look 
%like they could match but in fact don't.

%
\begin{figure}
	\makebox[0.5\textwidth][c]{%
		\begin{subfigure}[t]{0.048\textwidth}
			\centering
			\includegraphics[width=\textwidth]{images/pair_example1}
			\label{fig:fairey1}
		\end{subfigure}%
		\enspace %add desired spacing between images, e. g. ~, \quad, 
		%(or a blank line to force the subfigure onto a new line)
		\begin{subfigure}[t]{0.048\textwidth}
			\centering
			\includegraphics[width=\textwidth]{images/pair_example2}
			\label{fig:fairey2}
		\end{subfigure}%
		\enspace %
		\begin{subfigure}[t]{0.36\textwidth}
			\centering
			\includegraphics[width=\textwidth]{images/crop_examples}
		\end{subfigure}%
	}%
	\caption{Example of test set produced by a source pair of images}
	\label{fig:fairey}
\end{figure}
%
%\subsection{Experimental setup}
The \emph{MM} and \emph{MMC} algorithms have been tested against 
\emph{SIFT}, the algorithm Lowe proposed in \cite{lowe2004sift} as well 
as the \emph{Isodata} matching algorithm proposed in \cite{das2008event} 
which uses geometric information to improve the matching. The comparison 
has been done using the \emph{Murals} dataset, the \emph{Graf} set from 
\cite{mikolajczyk2005performance} and two images\footnote{100\_1942.jpg 
and 100\_1941.jpg} from \cite{gallagher2008} illustrated in figure 
\ref{fig:pitts_partition}.

Test sets have been generated from the image pairs by cropping out 
$250$px by $250$px patches with a random vertical and horizontal offset.  
Given a source image pair we produce $100$ pairs of patches where the 
two patches in any pair might or might not overlap.  This ensures that 
the patch pairs with no overlap still retain a basic similarity to each 
other, while the patches that do overlap often only overlap on a small 
part of their surface. Figure \ref{fig:fairey} shows an example of 
possible pairs of test images produced from a source image pair.  
Producing $n$ such pairs allows us to test not just how well the 
matching algorithm performs on a variety of overlap but also how many 
false positives we get on similar images that don't overlap at 
all\footnote{The set of source images and homographies as well as the 
	script to generate the cropped test sets based on them have been 
made available online together with the exact test sets used in this 
paper on 
\href{https://github.com/arnfred/Murals}{www.github.com/arnfred/Murals}}.

To judge whether a possible match ($p_1, p_2$) should be counted as an 
inlier given a homography $H$ relating two images $I_1$ and $I_2$ we 
test if $p_1$ and $p_2$ satisfies the following criterium:

$$\left\vert Hp_1 - p_2 \right\vert + \left\vert H^{-1}p_2 - p_1 
\right\vert < 5$$
%
\section{Results}
\label{results}

@@

\begin{figure}
	\centering
	\includegraphics[width=0.45\textwidth]{images/result_accumulated}
	\caption{Results on the \emph{Murals} dataset and the pair of img1 
		and img3 from the \emph{Graf} image set from 	
		\cite{mikolajczyk2005performance}. For each of the 9 source 
		image pairs, 100 test pairs have been created.  The x-axis 
	illustrates the accumulated returned matches for all pairs.}
	\label{fig:result_accumulated}
\end{figure}

\begin{figure}
	\centering
	\includegraphics[width=0.45\textwidth]{images/result_viewport}
	\caption{Results on viewport changes using the \emph{Graf} set from 
		\cite{mikolajczyk2005performance} as image pair sources. Img 1 
		to 6 in the \emph{Graph} set are images of the same subject but 
		with gradually increasing view port change. Img1 and 2 are 
		almost identical while img1 and 6 are taken from very different 
		angles. S is a testst of $100$ crops from img1 and img2. M is 
	likewise but based on img1 and img3.  L is img1 and img4 and XL is 
img1 and img5.  To decrease clutter \emph{MM} is not shown, but it 
performs alike to \emph{MMC}}
	\label{fig:result_viewport}
\end{figure}

Figure \ref{fig:result_graf} shows the results for 100 image pairs 
generated from img1 and img3 of the \emph{Graf} image set. The plot 
illustrates the results of \emph{SIFT}, \emph{MM}, \emph{MMC} and 
\emph{Isodata} as thresholds are varied over a range. \emph{Isodata} 
performs well with strict thresholds, but when we require more matches 
\emph{MM} and \emph{MMC} show superior results. To see if this results 
generalizes over more image sets, we tested the four algorithms on the 
\emph{Murals} dataset as well as the \emph{Graf} pair tested in fig.  
\ref{fig:result_graf}. In total 900 image pairs were generated from 9 
source pairs generating the plot shown in figure 
\ref{fig:result_accumulated} displaying the accumulated amount of 
matches on the x-axis. To verify that this performance is independent of 
the magnitude of view port change, we tested \emph{MMC} and \emph{SIFT} 
on the \emph{Graf} image set as shown in figure 
\ref{fig:result_viewport}. The figure confirms that \emph{MMC} is 
generally superior across view port changes.

For an example of a real life use case figure \ref{fig:result_pitts} 
shows the results on potentially overlapping image pairs generated from 
a typical holiday scene with partial overlap and slight view port 
change.

In terms of performance, table \ref{table:running_times} shows the 
running time of the four algorithms over $100$ image pairs of $250$px by 
$250$px. These numbers should be interpreted with the knowledge that 
much of the code behind \emph{MMC} and \emph{Isodata} is implemented in 
python while \emph{MM} and \emph{SIFT} makes use of opencv to execute 
computationally intensive operations in C++ which renders it much 
faster. While \emph{SIFT}, \emph{MM} and \emph{Isodata} can be 
implemented in $O(nlogn)$ where $n$ is the total amount of feature 
points, \emph{MMC} currently has a complexity of $O(n^2)$ due to the 
construction of a similarity matrix of the feature points. It is however 
possible to approximate this in $O(nlogn)$ using search trees, something 
that could be pursued in future work.
%
\begin{savenotes}
\begin{table}
	\centering
	\small
\begin{tabular}{l*{4}{c}}
	Algorithm: & \emph{SIFT} & \emph{MM} & \emph{MMC} & \emph{Isodata} 
	\\
	\noalign{\smallskip} 
	%
	Running time: & 20.94s & 18.23s & 883.99s & 555.33s \\
\end{tabular}
\caption{Running times as tested on a Intel(R) Core(TM) i5-3550 CPU @ 
3.30GHz with 8Gb memory}
\label{table:running_times}
\end{table}
\end{savenotes}
%

>>>>

\documentclass[conference]{IEEEtran}
%\documentclass{article}
\usepackage{footnote}
%\usepackage[font={small}]{caption}
\usepackage{sidecap}
\usepackage{graphicx}
\usepackage{cite}
\usepackage{hyperref}
\usepackage[stable]{footmisc}
\usepackage{caption}
\usepackage{adjustbox}
\usepackage{subcaption}
\usepackage{algpseudocode}
\usepackage{algorithm}
\usepackage{amsmath}
\usepackage{amssymb}

@@

where we do not know if two images have any corresponding areas or 

@@

constraints on the matches such as angular constraints 

@@

%proposes clustering feature points by using their geometric position 
%and exclude correspondences that are geometrically deviant. A more 
%sophisticated approach is introduced in \cite{wu2011robust} where the 
%areas are created by using the MSER feature detector to find areas of 
%interest in the images. These areas can be defined as an ellipse with a 
%certain radius around the interest points. The feature points that fall 
%within this ellipse are then matched according to epipolar constraints.

In practice however, there are many situations where using the geometry 
of the image to filter correspondences is not possible, such as object 
recognition where adjacent objects in one photo might be separated in 
another. Another example is images that do not overlap where algorithms 
might still find matches adhering to the imposed geometric constraints.
In addition methods using geometric constraints all require an initial 
set of correspondences. If this set of correspondences is narrowed down 
to a set with a higher ratio of correct matches the result of the 
geometric matching will benefit in terms of speed and accuracy. Finally 
some use cases might require a performance that can't be achieved by a 
more complex geometric method where simple non geometric methods might 
be able to provide sufficient matching accuracy with a smaller 
performance penalty.

@@

In turn the two methods proposed in this paper are inspired by a simple 
but novel idea. If we have two images and a given feature point in the 
first image is better matched with other feature points from the 
\emph{same} image than points in the other image, then any matches from 
this feature point to points in the other image is deemed unreliable and 
discarded.  This approach carries no implicit assumptions about the 
geometric consistency of matches and can as such easily be extended with 
other geometric solutions when appropriate.

@@

%compare two images that do not correspond. For any proposed 

@@

	\State $f_m,f_n \gets get2NNs(f_i, features \setminus 

@@

In figure \ref{fig:within} and \ref{fig:without} an example of the 
result of \emph{MM} is illustrated. In figure \ref{fig:unique} many of 
the incorrect matches occur in the fence in the bottom of the image.  
When we match all feature points together many of these incorrect 
matches are eliminated because points in the fence match with other 
points in the fence in the same image as shown in figure 
\ref{fig:within} and \ref{fig:without}.

\subsection{Mirror Match with Clustering (\emph{MMC})}
%
As opposed to \emph{MM}, \emph{MMC} diverges from traditional 
non-geometric feature matching by clustering the feature points by 
similarity. This process yields partitions of fairly similar feature 
points that we can match using the same approach as \emph{MM}. 
%
%Before we introduce the implementation details we will go over the 
%problem of graph clustering and how it relates to feature matching.
%
%\subsubsection{Graph Clustering}
%
%Given a set of feature points from two images, $im_1$ and $im_2$: $F_k = 
%\{k_i, k_j$ for $k_i \in im_1, k_j \in im_2\}$ as well as a matching 
%function $M(k_i, k_j) \rightarrow \mathbb{R}$ that takes two feature 
%points in an image and returns their matching score, we can define a 
%matrix $A$ where each element $A_ij = M(k_i, k_j)$. A can be interpreted 
%as the \emph{adjecency matrix} of the fully connected graph where each 
%vertex corresponds to a keypoint and the edge between two vertices has a 
%weight equal to the distance between the two corresponding keypoints.
%
%This representation reduces the problem of partitioning the keypoints 
%into groups to that of graph clustering or community structure depending 
%on the context. In the literature there are various ways of clustering a 
%graph according to different measures of what constitutes an optimal 
%partitioning. Traditionally the most used clustering algorithms have 
%been K-means and spectral clustering, but in recent years a host of new 
%algorithms have been proposed based on both Newman's concept of graph 
%modularity\footnote{Introduced in \cite{girvan2002}, discussed in 
%\cite{brandes2007} and used in \cite{blondel2008} as well as others} as 
%well as information theoretical measures\footnote{See for example 
%\cite{rosvall2008}} and the Potts spin model from physics\footnote{Used 
%in \cite{ronhovde2009}} just to mention a few approaches. Many of the 
%new algorithms differ from K-means clustering and Spectral clustering 
%in that they do not require the number of expected to clusters to be 
%specified beforehand\footnote{Among the aforementioned methods, this is 
%true for \cite{blondel2008} and \cite{rosvall2008}}.  Furthermore, on 
%tests done using randomly generated graphs with a known partitioning 
%\cite{blondel2008}, \cite{rosvall2008} and \cite{ronhovde2009} perform 
%markedly better than spectral clustering and 
%K-means\cite{lancichinetti2009}.
%
%The performance of clustering algorithms is a complicated issue, since 
%an optimal clustering given the same graph can vary depending on the 
%application. Spectral clustering for example will usually return a 
%partitioning where each partition is roughly equal in size\footnote{As 
%mentioned in \cite{von2007}} while the Louvain 
%clustering\cite{blondel2008} might return partitions of very uneven 
%size, even if the modularity measurement has been shown to penalize very 
%small clusters\cite{brandes2007}. Both behaviours can be beneficial 
%depending on the application, but when clustering feature points, 
%maintaining clusters of an even size usually means that some clusters 
%will be '\emph{catch-all}' clusters where the feature points that do 
%not fit anywhere else are grouped together. The necessity of specifying 
%the amount of partitions in for example Spectral clustering or Pott's 
%model clustering further exacerbates the issue since smaller partitions 
%are then combined into one to achieve the right amount of partitions.

@@

While the louvain clustering algorithm doesn't require any parameters in 
itself, it tends towards clustering all feature points together in the 
same partition if the graph is very connected.  To ensure that the graph 
is well clustered, the adjacency matrix is pruned so only edges above a 
certain threshold is kept. From empirical analysis keeping the 2.5\% of 
edges with the highest similarity seems to work well in practice. Figure 
\ref{fig:graph} shows the result of clustering the feature points 
displayed as a graph.
%
\begin{SCfigure*}
	\makebox[0.7\textwidth][c]{%
		\begin{subfigure}[t]{0.35\textwidth}
			\centering
			\includegraphics[width=\textwidth]{images/MMC_graph_full}
			\caption{Full Graph}
			\label{fig:full_graph}
		\end{subfigure}%
		~ %add desired spacing between images, e. g. ~, \quad, \qquad		  
		%(or a blank line to force the subfigure onto a new line)
		\begin{subfigure}[t]{0.32\textwidth}
			\centering
			\includegraphics[width=\textwidth]{images/MMC_graph_cropped}
			\caption{Cropped Section of Graph}
			\label{fig:cropped_graph}
		\end{subfigure}%
	}%
	\caption{The partitioned feature graph. Each vertex represents a 
		feature point while an edge represent a high similarity between 
		points. A partition is a connected group marked by a color.  Due 
		to a limited amount of colors some disconnected partitions have 
		been assigned the same color. The edge color of each node 
		signifies which image it belongs to. In the detailed crop to the 
		right the various cluster sizes can be seen, ranging from 
	hundreds of feature points to only two or three.}
	\label{fig:graph}
\end{SCfigure*}

@@

The matching algorithm for \emph{MMC}, \emph{getMatches}, finds matches 
within all partitions with more than two elements as done in \emph{MM}.  
However as can be seen in the example in figure \ref{fig:cropped_graph} 
many of the partitions contain only two feature points from different 
images linked by one edge. In the case of such a pair we compare the 
similarity of the these two feature points with their second best match 
and filters matches where this ratio falls over a certain threshold like 
in the \emph{SIFT} algorithm.

%\begin{algorithm}
%\caption{Impl. of getMatches (\emph{from MMC algorithm})}
%\label{alg-getmatches}
%{\fontsize{10}{10}\selectfont
%\begin{algorithmic}
%\Require $p$ : set of features, $t\in \mathbb{R}$, $features$ : Set of 
%all features
%\State $matches \gets \varnothing$
%\State $edges \gets \left\{similarity(f_i, f_j) \mid getImg(f_i) \neq 
%	getImg(f_j)$
%\State $\wedge f_i, f_j \in p \right\}$
%\If{$\left\vert edges \right\vert = 1$} \Comment If $\exists$ one edge 
%between images
%	\State $f_i,f_j \gets getFeatures(edges_{inter}, p)$
%	\State $f_m,f_n \gets get2NNs(f_i, features ~ 
%\setminus ~ \left\{f_i\right\})$
%	\State $f_s,f_t \gets get2NNs(f_j, features ~ 
%\setminus ~ \left\{f_j\right\})$
%	\State $ratio_i \gets similarity(f_i, f_n) / similarity(f_i, 
%f_j)$
%	\State $ratio_j \gets similarity(f_j, f_t) / similarity(f_j, 
%f_i)$
%	\If{$ratio_i < t\wedge ratio_j < t$}
%		\State $matches \gets matches \cup (f_i, f_j)$
%	\EndIf
%\ElsIf{$\left\vert edges \right\vert > 1$} \Comment If $\exists$ more 
%than one edge between images
%	\ForAll{$f_i \in p$}
%		\State $f_m,f_n \gets get2NNs(f_i, p \setminus 
%\left\{f_i\right\})$
%		\State $ratio \gets similarity(f_i, f_n) / similarity(f_i, 
%f_m)$
%		\If{$getImg(f_i) \neq getImg(f_m) \wedge ratio < t
%\wedge (f_m, f_i) \not\in matches$}
%			\State $matches \gets matches \cup (f_i, f_m)$
%		\EndIf
%	\EndFor
%\EndIf
%
%\Return matches
%\end{algorithmic}
%}
%\end{algorithm}

\section{Experiments}
\label{experiment}
%
To reliably measure the accuracy of a matching method on real images we 
either need a set of image pairs tied by a homography or manually count 
the number of inliers. The latter becomes unfeasible the moment we 
attempt to test a non-trivial set of images.

In \cite{mikolajczyk2005performance} Mikolajczyk and Schmid introduces a 
set of images to compare the performance of feature detectors. The test 
set covers different types of image variance such as different lighting, 
blur, rotation as well as viewpoint change. Inspired by this 
dataset\footnote{In particular by the 'Graf' image set} and motivated by 
the need for more image pairs featuring viewport changes, we have 
compiled a set of $8$ image pairs consisting of subjects taken from two 
different angles. The images are collected from Flickr's database of 
images published under a creative commons 
licence\footnote{\href{http://creativecommons.org/}{www.creativecommons.org}, 
a copyleft license permitting redistribution} and feature murals which 
makes it possible to relate points in the image pairs with a homography.  
The images have been cropped to feature the same motive and resized to 
$900$px by $600$px. This dataset will be referred to as the 
\emph{Murals} dataset in this paper. The images can be seen in figure 
\ref{fig:murals}.

@@

%Inspired by the graffiti image we have collected a set of image pairs 
%featuring murals of a few different artists\footnote{Including Banksy, 
%Blu and Shepard Fairey among others}. Each pair consist of images of 
%the same motive taken from different angles, often including repetitive 
%patterns and texture. Based on this image set the two algorithms are 
%tested on a series of image patches from each pair. To verify that the 
%algorithms proposed are reliable we need to test that we get good 
%matches when images overlap and that we get no matches when they do 
%not.  This means that it isn't enough just testing against pairs of 
%images that we know match.  We also need to test against pairs that 
%might look like they could match but in fact do not.

%
\begin{figure}
	\makebox[0.5\textwidth][c]{%
		\begin{subfigure}[t]{0.048\textwidth}
			\centering
			\includegraphics[width=\textwidth]{images/pair_example1}
			\label{fig:fairey1}
		\end{subfigure}%
		\enspace %add desired spacing between images, e. g. ~, \quad, 
		%(or a blank line to force the subfigure onto a new line)
		\begin{subfigure}[t]{0.048\textwidth}
			\centering
			\includegraphics[width=\textwidth]{images/pair_example2}
			\label{fig:fairey2}
		\end{subfigure}%
		\enspace %
		\begin{subfigure}[t]{0.36\textwidth}
			\centering
			\includegraphics[width=\textwidth]{images/crop_examples}
		\end{subfigure}%
	}%
	\caption{Example of test set produced by a source pair of images}
	\label{fig:fairey}
\end{figure}
%
%\subsection{Experimental setup}
The \emph{MM} and \emph{MMC} algorithms have been tested against 
\emph{SIFT}, the algorithm Lowe proposed in \cite{lowe2004sift} as well 
as the \emph{Isodata} matching algorithm proposed in \cite{das2008event} 
which uses geometric information to improve the matching. The comparison 
has been done using the \emph{Murals} dataset, the \emph{Graf} set from 
\cite{mikolajczyk2005performance} and two images\footnote{100\_1942.jpg 
and 100\_1941.jpg} from \cite{gallagher2008} illustrated in figure 
\ref{fig:comparemirror}.

Test sets have been generated from the image pairs by cropping out 
$250$px by $250$px patches with a random vertical and horizontal offset.  
Given a source image pair we produce $100$ pairs of patches where the 
two patches in any pair might or might not overlap.  This ensures that 
the patch pairs with no overlap still retain a basic similarity to each 
other, while the patches that do overlap often only overlap on a small 
part of their surface. Figure \ref{fig:fairey} shows an example of 
possible pairs of test images produced from a source image pair.  
Producing $n$ such pairs allows us to test not just how well the 
matching algorithm performs on a variety of overlap but also how many 
false positives we get on similar images that do not overlap at 
all\footnote{The set of source images and homographies as well as the 
	script to generate the cropped test sets based on them have been 
	made available online together with the exact test sets used in this 
paper on 
\href{https://github.com/arnfred/Murals}{www.github.com/arnfred/Murals}}.  
In practice the amount of overlap between pairs in a test set will 
depend on the overlap and viewport change of the source image pair.  To 
give a rough impression of a typical example, table \ref{table:overlap} 
shows the overlap of image pairs created from img1 and img3 of the 
\emph{Graf} image set from \cite{mikolajczyk2005performance}.

Given a potential match between two images $I_1$ and $I_2$, $m = 
\left(p_1, p_2\right)$ and a homography $H$ relating $I_1$ to $I_2$, we 
can calculate if $m$ is an inlier by check if the positions $p_1$ and 
$p_2$ satisfies the following criteria:
$$\left\vert Hp_1 - p_2 \right\vert + \left\vert H^{-1}p_2 - p_1 
\right\vert < 5$$
That is, the distance between $p_1$ translated to $I_2$ and $p_2$ 
\emph{plus} the distance between $p_2$ translated to $I_1$ should be 
less than five pixels.

\begin{table}
	\centering
	\small
\begin{tabular}{l*{3}{r}}
	Overlap: & 0\% & $< 50$\% & $> 50$\%  \\
	\noalign{\smallskip}
	%
	Percent of image pairs: & 21\% & 54\% & 25\% \\
\end{tabular}
\caption{Overlap on the testset of 100 image pairs created from img1 and 
img3 of the \emph{Graf} image set. 21 images had no overlap}
\label{table:overlap}
\end{table}
\section{Results}
\label{results}

@@

\begin{figure}
	\centering
	\includegraphics[width=0.5\textwidth]{images/result_accumulated}
	\caption{Results on the \emph{Murals} dataset and the pair of img1 
		and img3 from the \emph{Graf} image set from 	
		\cite{mikolajczyk2005performance}. For each of the 9 source 
		image pairs, 100 test pairs have been created.  The x-axis 
	illustrates the accumulated returned matches for all pairs.}
	\label{fig:result_accumulated}
\end{figure}

\begin{figure}
	\centering
	\includegraphics[width=0.45\textwidth]{images/result_viewport}
	\caption{Results on viewport changes using the \emph{Graf} set from 
		\cite{mikolajczyk2005performance} as image pair sources. Img 1 
		to 6 in the \emph{Graph} set are images of the same subject but 
		with gradually increasing viewport change. Img1 and 2 are 
		almost identical while img1 and 6 are taken from very different 
		angles. S is a testset of $100$ crops from img1 and img2. M is 
	likewise but based on img1 and img3.  L is img1 and img4 and XL is 
img1 and img5.  To decrease clutter \emph{MM} is not shown, but it 
performs alike to \emph{MMC}}
	\label{fig:result_viewport}
\end{figure}

Figure \ref{fig:result_graf} shows the results for 100 image pairs 
generated from img1 and img3 of the \emph{Graf} image set. The plot 
illustrates the results of \emph{SIFT}, \emph{MM}, \emph{MMC} and 
\emph{Isodata} as thresholds are varied over a range. \emph{Isodata} 
performs well with strict thresholds, but when we require more matches 
\emph{MM} and \emph{MMC} show superior results. To see if this results 
generalizes over more image sets, we tested the four algorithms on the 
\emph{Murals} dataset as well as the \emph{Graf} pair tested in fig.  
\ref{fig:result_graf}. In total 900 image pairs were generated from 9 
source pairs generating the plot shown in figure 
\ref{fig:result_accumulated} displaying the accumulated amount of 
matches on the x-axis. To verify that this performance is independent of 
the magnitude of viewport change, we tested \emph{MMC} and \emph{SIFT} 
on the \emph{Graf} image set as shown in figure 
\ref{fig:result_viewport}. The figure confirms that \emph{MMC} is 
generally superior across viewport changes.

For an example of a real life use case, figure \ref{fig:result_pitts} 
shows the results on 100 potentially overlapping image pairs generated 
from a typical holiday scene featuring occlusion and a slight viewport 
change.

In terms of performance, table \ref{table:running_times} shows the 
running time of the four algorithms over $100$ image pairs of $250$px by 
$250$px. These numbers should be interpreted with the knowledge that 
much of the code behind \emph{MMC} and \emph{Isodata} is implemented in 
python while \emph{MM} and \emph{SIFT} makes use of opencv to execute 
computationally intensive operations in C++ which renders it much 
faster. While \emph{SIFT}, \emph{MM} and \emph{Isodata} can be 
implemented in $O(nlogn)$ where $n$ is the total amount of feature 
points, \emph{MMC} currently has a complexity of $O(n^2)$ due to the 
construction of a similarity matrix of the feature points. It is however 
possible to approximate this in $O(nlogn)$ using search trees, something 
that could be pursued in future work.
%
\begin{table}
	\centering
	\small
\begin{tabular}{l*{4}{c}}
	Algorithm: & \emph{SIFT} & \emph{MM} & \emph{MMC} & \emph{Isodata} 
	\\
	\noalign{\smallskip} 
	%
	Running time: & 20.94s & 18.23s & 883.99s & 555.33s \\
\end{tabular}
\caption{Running times as tested on a Intel(R) Core(TM) i5-3550 CPU @ 
3.30GHz with 8Gb memory}
\label{table:running_times}
\end{table}
%
